---
title: "CLV Exploratory"
author: "Josh Kong"
date: "10/13/2020"
output: pdf_document
editor_options: 
  chunk_output_type: console
---

```{r}
library(tidyverse)
library(lubridate)
retail <- read_csv("retail_cleaned.csv")
theme_light()
```

## Using skimr to get a gist for how the dataset looks

```{r}
skimr::skim(retail)
```

The only data that we appear to be missing is CustomerID. This is probably due to the customer only making a one time purchase without making an account. I plan to filter these customers out later when I predict the customer's future CLV values.

```{r}
range(retail$InvoiceDate)
#the data only accounts for a month and half of data.
```


```{r}
#looking at revenue by the dates
revenue_by_month <- retail %>% 
  group_by(month = floor_date(InvoiceDate, "month")) %>% 
  summarise(revenue = sum(sales)) %>% 
  ungroup()

 revenue_by_month %>% 
  ggplot(aes(month, revenue)) + 
  geom_point()+
  geom_line() +
   labs(x = "", y = "Revenue", title = "Revenue by Month")

#there is a very large spike in revenue during November. This is probably due to Black Friday sales, or early Christmas shopping.
#december 2011 has low numbers in sales. this is most likely due to the data not being complete
```


```{r}
#looking at the monthly growth rate
monthly_growth <- rep(NA, nrow(revenue_by_month))
monthly_growth[1] <- 0
for(i in 2:nrow(revenue_by_month-1))
{
  monthly_growth[i] <- (revenue_by_month$revenue[i] - revenue_by_month$revenue[i-1])/revenue_by_month$revenue[i-1]
}
revenue_by_month$monthly_growth <- monthly_growth
revenue_by_month %>% 
  filter(month != "2011-12-01") %>% 
  ggplot(aes(month, monthly_growth)) + 
  geom_point() + 
  geom_line()+
  geom_hline(yintercept = 0, lty = 2) + 
  labs(title = "Monthly Growth", y = "") + 
  scale_y_continuous(labels = scales::percent_format())
#I'm excluding 2011 December from this graph because the data is still incomplete.
```

```{r}
#looking at monthly active customers
retail %>% 
  filter(!is.na(CustomerID)) %>% 
  group_by(month = floor_date(InvoiceDate, "month"), CustomerID) %>% 
  summarise(count = n()) %>% 
  ungroup() %>% 
  count(month) %>% 
  filter(month != "2011-12-01") %>% 
  ggplot(aes(month, n)) + 
  geom_point() + 
  geom_line()+ 
  labs(y = "Number of Active Users", x = "", title = "Number of Active Users per Month")

#looks like there is consistent growth in the number of active users per month
```

```{r}
#taking a look a new customers per month
retail %>% 
  distinct(CustomerID, InvoiceDate) %>% 
  group_by(CustomerID) %>% 
  summarise(first_purchase_date = min(InvoiceDate)) %>% 
  ungroup() %>% 
  count(first_purchase_date) %>% 
  group_by(month = floor_date(first_purchase_date, "month")) %>% 
  summarise(new_customers = sum(n)) %>% 
  filter(month != "2011-12-01", month != "2010-12-01") %>%  #filtering out december because data is still incomplete
  ggplot(aes(month, new_customers)) + 
  geom_point() + 
  geom_line()
#we didn't have many new customers from June through August. Starting September there was a spike in new customers once again
```

## RFM (recency, frequency, monetary value) Clustering
```{r}
library(cluster)
library(factoextra)
```

Calculating recency. Recency score is equal to the most recent purchase subtracted by the customer's most recent purchase.

```{r}
#going to calculate recency by subtracting the customer's most recent transcation by the max of the invoice date
maxdate <- max(retail$InvoiceDate)

recency_scores <- retail %>% 
  filter(!is.na(CustomerID)) %>% 
  group_by(CustomerID) %>% 
  summarise(recency = maxdate - max(InvoiceDate)) 

#distribution of recency scores
recency_scores %>% 
  ggplot(aes(recency)) + 
  geom_histogram()
```

Choosing the number of clusters for recency and then using kmeans clustering
```{r}
#elbow method to see optimal number of clusters
fviz_nbclust(as.matrix(recency_scores$recency), kmeans, method = "wss")

#looking at the elbow plot, it seems it would be optimal to do 4 clusters

recency_cluster <- kmeans(recency_scores$recency, 4)
recency_center <- sort(desc(recency_cluster$centers)) * -1
recency_cluster <- kmeans(recency_scores$recency, centers = recency_center) 
#doing it twice to order the clusters by the cluster means
recency_scores$r_clus <- recency_cluster$cluster

#the higher the score, the more recently they've purchased an item
```

Calculating Frequency. Total number of order by a certain customer.

```{r}
#We'll be counting each invoice number as one order.
frequency_scores <- retail %>% 
  filter(!is.na(CustomerID)) %>% 
  count(CustomerID, InvoiceNo) %>% 
  count(CustomerID, sort = TRUE)

#distribution of the frequency scores
retail %>% 
  filter(!is.na(CustomerID)) %>% 
  count(CustomerID, InvoiceNo) %>% 
  count(CustomerID, sort = TRUE) %>% 
  ggplot(aes(n)) + 
  geom_histogram(binwidth = 15)
```

Choosing the optimal number of clusters for frequency and doing k means clustering.

```{r}
fviz_nbclust(as.matrix(frequency_scores$n), kmeans, method = "wss")

#looking at the elbow plot, it seems it would be optimal to do 5 clusters. However, to be consistent with recency, I will do 4 clusters once again.

frequency_cluster <- kmeans(frequency_scores$n, 4)
frequency_center <- sort(frequency_cluster$centers)
frequency_cluster <- kmeans(frequency_scores$n, centers = frequency_center) 
#doing it twice to order the clusters by the cluster means
frequency_scores$f_clus<- frequency_cluster$cluster
```


Calculating monetary value
```{r}
#we are going to be calculating the total revenue each customer brought in
money_scores <- retail %>% 
  filter(!is.na(CustomerID)) %>% 
  group_by(CustomerID) %>% 
  summarise(total_revenue = sum(sales))

#looking at the distribution of monetary value
money_scores %>% 
  ggplot(aes(total_revenue)) + 
  geom_histogram()
```

Choosing optimal number of clusters for monetary value and doing k means clustering
```{r}
fviz_nbclust(as.matrix(money_scores$total_revenue), kmeans, method = "wss")

#looking at the elbow plot, it seems it would be optimal to do 4 clusters. 

money_cluster <- kmeans(money_scores$total_revenue, 4)
money_center <- sort(money_cluster$centers)
money_cluster <- kmeans(money_scores$total_revenue, centers = money_center) 
#doing it twice to order the clusters by the cluster means
money_scores$m_clus <- money_cluster$cluster
```

Combining all the scores together
```{r}
frequency_scores <- frequency_scores %>% 
  arrange(CustomerID)

rfm_scores <- as_tibble(data.frame(recency_scores, frequency_scores[-1], money_scores[-1]))
colnames(rfm_scores) <- c("CustomerID","recency","r_clus","frequency","f_clus","money_value","m_clus")
rfm_scores <- rfm_scores %>% 
  group_by(CustomerID) %>% 
  mutate(score = sum(r_clus,f_clus,m_clus)) %>% 
  ungroup()


#looking at the distribution of scores. Nobody got a perfect score of 12
rfm_scores %>% 
  count(score) %>% 
  ggplot(aes(as.factor(score), n)) + 
  geom_col() +
  labs(title= "Distribution of Scores", x = "Score", y= "")
```

The scores range from 3-12. 

I will value 3-4 as low value customers.
5-7 as mid value customers.
8-11 as high value customers.

```{r}
for(i in 1:nrow(rfm_scores))
{
  if(rfm_scores$score[i] <= 4)
  {
    rfm_scores$value[i] = "low"
  }
  else if(rfm_scores$score[i] > 4 && rfm_scores$score[i] <8)
  {
    rfm_scores$value[i] = "mid"
  }
  else
  {
    rfm_scores$value[i] = "high"
  }
}
rfm_scores$value <- as.factor(rfm_scores$value)
```

Taking a look at how our valued customers are distributed in a scatterplot.
```{r}
rfm_scores %>% 
  ggplot(aes(recency, frequency, color = value)) +
  geom_point()

rfm_scores %>% 
  ggplot(aes(recency, money_value, color = value))+
  geom_point()

rfm_scores %>% 
  ggplot(aes(frequency,money_value, color = value)) + 
  geom_point()
```


Saving the scores dataset
```{r}
write.csv(rfm_scores, "rfm_scores.csv")
```










